---
title: "ami_spinUp_howto"
author: "Lincoln Harris"
date: "4/16/2018"
output: html_document
---

*Instructions for how to to set up and run programs on an Ubuntu AMI that has been pre-configured with a number of useful programs including STAR, Trinity & BLAST.*      

## Configure machine

**From your AWS EC2 console:** 
  1. Make sure region is *US West (Oregon)*
  2. Go to *IMAGES* > *AMIs*
  3. Enter *marvin* in the search bar
  4. ctrl-click on this AMI, and select *launch* from the drop down menu  
  5. Choose an instance type. For STAR index generate, I usually select m4.10xlarge
  6. Select *Review & Launch*            
  7. Select *Launch*          
  8. Select your key pair (.pem) from the drop down menu
  9. Select *Launch instance*         
  10. Select *View Instances*         
  11. From the display screen of running instances, search for yours by the name of you key pair (it sometimes takes a minute before you can view your instance)
  12. Once you've located your instance, give it a descriptive name by clicking on the name field
  13. As soon as the *Status Check* field for your instance moves from *Initializing* to *2/2 checks*, you're ready to login

To login to your VM: 
```{r}
ssh -i /path/to/keyPair/myFile.pem ubuntu@<PublicDNS>
```
you can access the Public DNS (IPv4) for your instance on your AWS console

## Push up your data

This will copy a file from your local machine to the home (~) directory of your VM **(From local terminal)**                     
```{r}
scp -i /path/to/keyPair/myFile.pem /path/to/local/file ubuntu@<PublicDNS>:~ 
```

## Pull down your data

To copy a single file **(From local terminal)**                       
```{r}
scp -i /path/to/keyPair/myFile.pem ubuntu@<PublicDNS>:~/path/to/file/on/VM /destination/on/local
```

or an entire directory **(From local terminal)**         
```{r}
scp -i /path/to/keyPair/myFile.pem -r ubuntu@<PublicDNS>:~/path/to/dir/on/VM /destination/on/local
```

## Shutting Down

To exit the instance
```{r}
logout
```
(sometimes need to enter this twice)

Back in the console, ctrl-click the instance name, and select *terminate*   
