---
title: "rstudio_VM_howto.Rmd"
author: "Lincoln Harris"
date: "7/2/2018"
output: html_document
---

*Instructions for how to set up and run RStudio on a Ubuntu AMI*

## Configure your VM

**From your EC2 console:**                 
  1. Make sure region is Oregon    
  2. Select *Launch Instance*          
  3. Select *Community AMIs* from the left most panel   
  4. In the search bar, enter *rstudio julia*       
  5. Select the AMI you want to spin up              
      (I usually go with *RStudio-1.0.153_R-3.4.1_Julia-0.6.0_ubuntu-16.04-LTS-64bit*)        
  6. Choose the instance type you want            
      (I generally go with m4.4xlarge when I have to do any serious computation)    
          *Review & Launch*       
  7. Once on the next screen, select *Edit security groups*       
  8. Click *Select an existing security group*      
  9. Scroll down until you find *lincoln-rstudio*       
          *Review & Launch*      
  10. Select *Launch*   
  11. Select your key-pair from the drop down menu          
          *Launch Instances*    
  12. From the green dialog box, click on your instance and be sure to give it a descriptive name     
      Your instance is now spinning up. As soon as you see *2/2 checks* in the **Status Checks** column you'll be ready to work with it             
  13. With you instance selected, grab the *IPv4 Public IP* and paste it into your browser
  14. Sign in -- Default user name and password are both *rstudio*    
  15. Follow the prompt in selecting a more secure password. You are now ready to work on your virtual Rstudio server!         

## Push up your data

This will copy a file from your local machine to the home (~) directory of your VM **(From local terminal)**                     
```{r}
scp -i /path/to/keyPair/myFile.pem /path/to/local/file ubuntu@<PublicDNS>:~ 
```

After this you will need to move the file from ~ to a directory that the Rstudio server recognizes. 
To do this we first need to ssh into the virtual machine

```{r}
ssh -i /path/to/keyPair/myFile.pem ubuntu@<PublicDNS>
```

Once you're logged in, you should be able to see the file you just pushed up. To move it to the right place: 
```{r}
cp myFile /home/rstudio
```

You should now be able to view the file from your Rstudio web console. 

## Pull down your data

To copy a single file **(From local terminal)**                       
```{r}
scp -i /path/to/keyPair/myFile.pem ubuntu@<PublicDNS>:~/path/to/file/on/VM /destination/on/local
```

or an entire directory **(From local terminal)**         
```{r}
scp -i /path/to/keyPair/myFile.pem -r ubuntu@<PublicDNS>:~/path/to/dir/on/VM /destination/on/local
```

Keep in mind that you can also save image files (pdfs, pngs, etc.) like you normally would with Rstudio, and then open them with your web browser and save them directly to your local machine. 

## Shutting Down
To exit your terminal open on the VM:
```{r}
logout
```
(sometimes need to enter this twice)

Back in the console, *ctrl-click* the instance name, and select *stop* 

Selecting *stop* allows you to shut down the VM without killing it, which means you'll preserve whatever data you've pushed up there and any system modifications you've made. 

Also keep in mind that once you've stopped an instance, you can change the instance type. **This allows you to spin up something bigger without recreating the instance or repushing your data!!**

