---
title: "STAR_VM_install_and_run.Rmd"
author: "Lincoln Harris"
date: "3/9/2018"
output: html_document
---

*Instructions for how to to run STAR on an Ubuntu AMI*   

## Configure your VM

**From your AWS console:**                 
   1. Make sure region is N.Virginia 
    2. Go to *IMAGES* > *AMIs*        
    3. Enter *linc_image_gen_t3* in the search bar
    4. ctrl-click on this AMI, and select *launch* from the drop down menu      
        **This AMI is pre-configured with STAR_2.5.3a_modified, and 500 GiB physical storage**     
    5. Choose an instance type. For STAR index generate, I usually select m4.10xlarge
    6. Select *Review & Launch*            
    7. Select *Launch*          
    8. Select your key pair (.pem) from the drop down menu
    9. Select *Launch instance*         
    10. Select *View Instances*         
    11. From the display screen of running instances, search for yours by the name of you key pair (it sometimes takes a minute before you can view your instance)
    12. Once you've located your instance, give it a descriptive name by clicking on the name field
    13. As soon as the *Status Check* field for your instance moves from *Initializing* to *2/2 checks*, you're ready to login
  
  
To login to your VM: 
```{r}
ssh -i /path/to/keyPair/myFile.pem ubuntu@<PublicDNS>
```
you can access the Public DNS (IPv4) for your instance on your AWS console

Run just to make sure everything is working
```{r}
STAR --version
```

## Push up your data

This will copy a file from your local machine to the home (~) directory of your VM **(From local terminal)**                     
```{r}
scp -i /path/to/keyPair/myFile.pem /path/to/local/file ubuntu@<PublicDNS>:~ 
```

To run STAR you'll need to push up **.fa** & **.gtf** files for your genome

## Run STAR

**Step 1 -- Generate genome index**         

we'll run STAR from within a new directory 
```{r}
mkdir wkdir
cd wkdir
```

copy files into working directory
```{r}
cp ../*.fa ./
cp ../*.gtf ./
```
   
Create a directory where the genome indicies will be stored
```{r}
mkdir my_genome_dir
```
 
Give STAR special permissions to write to this dir
```{r}
sudo chmod -R 777 my_genome_dir/
```

Run STAR genome generate  
```{r}
STAR --runThreadN 32 --runMode genomeGenerate --genomeDir ./my_genome_dir/ --genomeFastaFiles ./*.fa --sjdbGTFfile ./*.gtf --sjdbOverhang 75
```

The genome index will be written to my_genome_dir/

**Step 2 -- Run Mapping**           

These are the params I run mapping with -- pretty sure they're exactly the same as the aegea STAR pipeline
```{r}
STAR --runThreadN 12 --genomeDir genomeDir/ --readFilesIn ./A12-1000500501/A12-1000500501_R1.fasta ./A12-1000500501/A12-1000500501_R2.fasta --twopassMode Basic --outSJfilterCountTotalMin 1 1 1 1 --outSJfilterCountUniqueMin 1 1 1 1 --outFilterMultimapNmax 20 --alignSJoverhangMin 8 --alignSJDBoverhangMin 1 --outFilterMismatchNmax 999 --outFilterMismatchNoverLmax 0.04 --alignIntronMin 20 --alignIntronMax 1000000 --alignMatesGapMax 1000000 --outSAMstrandField intronMotif
```

## Pull down your data

To copy a single file **(From local terminal)**                       
```{r}
scp -i /path/to/keyPair/myFile.pem ubuntu@<PublicDNS>:~/path/to/file/on/VM /destination/on/local
```

or an entire directory **(From local terminal)**         
```{r}
scp -i /path/to/keyPair/myFile.pem -r ubuntu@<PublicDNS>:~/path/to/dir/on/VM /destination/on/local
```

## Shutting Down

To exit the instance
```{r}
logout
```
(sometimes need to enter this twice)

Back in the console, ctrl-click the instance name, and select *terminate*     

